Learning constraints in spreadsheets and tabular data

### 2016-09-23 Paper/ECAI/tabular.tex
Edits evaluation and constraints section (fail fast)

\author{Samuel Kolb \and Sergey Paramonov \and Tias Guns \and Luc De Raedt}
\institute{KULeuven}

@@

For \textit{FOREIGNKEY}, \textit{LOOKUP}, aggregates and \textit{PROJECT} we use custom, optimized methods to find subassignments.
Instead of the more generic generate-and-test implementation, we use, for example, caching, maximal range-finding and partial tests.
While conditional aggregates use generate-and-test, they employ similar optimizations in their test method.
Based on the assumptions that most subassignments will not satisfy a constraint template, the implementations are geared to fail-fast when possible.

@@

All spreadsheets also have a set of ground-truth of constraints, we call these the \textbf{intended} constraints, that are expected to be (re-)discovered.
These were determined by us using either the formulas of the original sheet (if present), or inferring them based on, for example, the header names in the sheet.
In the following experiments, we focus on spreadsheets specifically and hence only include \textit{functional} constraints, constraints that can be expressed as formulas. We ignoring structural constraints such as \textit{ALLDIFFERENT}, \textit{FOREIGNKEY} and \textit{PERMUTATION}. All constraints are stored in their canonical form. 5 spreadsheets contained rows or columns that were entirely identical, and any intended constraint involving one such vector was also added for the equivalent vectors.

@@

\paragraph{Theoretical analysis of runtime, this does not belong in experiments}\tias{move to technical section}\sam{Has to be moved yet to 'Constraints' section, explaining fail-fast}
For most constraint templates the time required to check whether a subassignment satisfies the template is linear in the size of the subassignment (the number of subblocks and their size).
For a constraint with arity~$n$, and a subassignment consisting of $n$~blocks with $m$~vectors of length~$k$, the worst case complexity becomes $O(n * m * k)$.
However, in practice, most subassignments will not satisfy the constraint template.
Since our implementation tries to account for this by failing fast\tias{explain more precisely what you mean with 'failing fast'}, the amortized time is constant in the length of the vectors~$k$, unless $k$~is very large\footnote{Specifically, if $k$ is much larger than the number of candidate subassignments} or many constraints hold in the data.
For templates requiring their arguments to be single vectors $m$~is~$1$, and the total complexity for checking a subassignment becomes~$O(n) \simeq O(1)$ (since $n$ is typically small).

@@

\paragraph{Q2: Intended and supported constraints}
%\item Q2: it may fail to find intended constraints because it does not support such constraints; how does or method perform on real spreadsheets?
The overall recall achieved by \sname is~$0.88~(81/92)$. The overall and average recall per category is also high, as shown in Table~\ref{tbl:category_overview}.
The intended constraints not discovered correspond to constraints not supported by the system (supported recall is 1). \tias{Rewrite and go into more detail about all supported found, some intended not support, which ones, what category, just some more details}

\paragraph{Q3: precision}
%\item Q3: it may find non-intended constraints; how many and what type of non-intended constraints are found on real spreadsheets?
The final question, Q3, is about the amount and type of non-intended constraints found.
Across all spreadsheets \sname achieves a precision of~$0.79~(81/102)$.
While the average precision per spreadsheet is over $0.90$ for all categories, the overall precision is much higher for the exercises and data categories ($0.94$ and $1.00$) than for the tutorials category ($0.70$).

Examining the tutorial spreadsheets confirms that there are a few spreadsheets that have a high number of additional constraints. This is caused by duplicate data or multiple ways to calculate the same vectors. \tias{More details or examples; we copy intended constraints for equivalent/duplicate vectors, so not an issue?}
For these spreadsheets the precision could be increased by post-processing the constraints to detect implications or heuristically selecting one among multiple ways to compute the same vector.

>>>>

\author{Samuel Kolb* \and Sergey Paramonov* \and Tias Guns \and Luc De Raedt}
\institute{KULeuven\\ {*} Shared first author}

@@

For \textit{FOREIGNKEY}, \textit{LOOKUP}, aggregates and \textit{PROJECT} we use custom, optimized methods to find subassignments.
Instead of the more generic generate-and-test implementation, we use, for example, caching, maximal range-finding and partial tests.
While conditional aggregates use generate-and-test, they employ similar optimizations in their test method.

Based on the assumptions that most subassignments will not satisfy a constraint template, the implementations are geared to \textit{fail-fast} when possible.
Specifically, the tests try to prove that a subassignment does not satisfy a constraint template as fast as possible.
For example, it can be often shown that a constraint does not hold by only looking first entries of every subblock (e.g. if $\sbl{r}[0] \neq \sbl{1}[0] \times \sbl{2}[0]$ then $\sbl{r} = \sbl{1} \times \sbl{2}$ does not hold).
As a result, the time to check if a subassignment satisfies a constraint templates with a fail-fast implementation requires constant time\footnote{If the number of entries to look at is bounded by a constant, the vector lengths are smaller than the number of subassignments to test ($a$) and the time to test subassignments that satisfy the constraint is $O(a)$, then the time to test a subassignment is amortized $O(1)$} in most cases and does not depend on the length of the vectors.

%\paragraph{Theoretical analysis of runtime, this does not belong in experiments}\tias{move to technical section}\sam{Has to be moved yet to 'Constraints' section, explaining fail-fast}
%For most constraint templates the time required to check whether a subassignment satisfies the template is linear in the size of the subassignment (the number of subblocks and their size).
%For a constraint with arity~$n$, and a subassignment consisting of $n$~blocks with $m$~vectors of length~$k$, the worst case complexity becomes $O(n * m * k)$.
%However, in practice, most subassignments will not satisfy the constraint template.
%Since our implementation tries to account for this by failing fast\tias{explain more precisely what you mean with 'failing fast'}, the amortized time is constant in the length of the vectors~$k$, unless $k$~is very large\footnote{Specifically, if $k$ is much larger than the number of candidate subassignments} or many constraints hold in the data.
%For templates requiring their arguments to be single vectors $m$~is~$1$, and the total complexity for checking a subassignment becomes~$O(n) \simeq O(1)$ (since $n$ is typically small).

@@

All spreadsheets also have a set of ground-truth of constraints, we call these the \textbf{intended} constraints, that are expected to be (re-)discovered.
These were determined by us using either the formulas of the original sheet (if present), or inferring them based on, for example, the header names in the sheet.
In the following experiments, we focus on spreadsheets specifically and hence only include \textit{functional} constraints, constraints that can be expressed as formulas. We ignoring structural constraints such as \textit{ALLDIFFERENT}, \textit{FOREIGNKEY} and \textit{PERMUTATION}. All constraints are stored in their canonical form. 5 spreadsheets contained rows or columns that were entirely identical, and any intended constraint involving one such vector was also added for the equivalent vectors. \samuel{They are usually not copied}

@@

@@

\paragraph{Q2: Intended and supported constraints}
%\item Q2: it may fail to find intended constraints because it does not support such constraints; how does or method perform on real spreadsheets?
The overall recall achieved by \sname is~$0.88~(81/92)$.
When looking at the recall per category, \sname is able to consistently obtain high values for both overall and average recall, as shown in Table~\ref{tbl:category_overview}.
The intended constraints not discovered correspond to constraints that are not supported by the system (supported recall is 1).
Therefore, \sname is able to find all constraints for supported constraint templates, which confirms that our bias (signature) is not too restrictive.

The intended constraints that are not discovered correspond with nested constraints (e.g. $R = \frac{W}{(L/100)^2}$), variants of supported constraints (e.g. series that do not start at 1) or more complex versions of supported constraints (e.g. conditional aggregates that automatically infer the condition).

\paragraph{Q3: Precision}
%\item Q3: it may find non-intended constraints; how many and what type of non-intended constraints are found on real spreadsheets?
The final question, Q3, is about the amount and type of non-intended constraints found.
Across all spreadsheets \sname achieves a precision of~$0.79~(81/103)$.
While the average precision per spreadsheet is high, over $0.90$ in every category, however the overall precision is much higher for the exercises and data categories ($0.94$ and $1.00$) than for the tutorials category ($0.70$).

Examining the tutorial spreadsheets confirms that there are a few spreadsheets that have a high number of additional constraints, e.g. an spreadsheet computes aggregates on its inventory but copies the data column for every aggregate.
Since every aggregate can be calculated based on any of the (equal) columns, many of the resulting constraints will be marked as unintended.
Usually, unintended constraints are the result of multiple ways to calculate the same result, e.g. if there is duplicate data.
A way to increase the precision is by post-processing the constraints to detect implications and heuristically selecting one among multiple ways to compute the same vector.
